Practical Concurrent Priority Queues \\ Student Paper for Seminar in Algorithms 2013W \\ Technical University of Vienna

### 2014-04-04 paper/paper.Rnw
Add section on synchronization primitives

@@

An area in memory accessed frequently by a large number of processes is said to be \emph{contended}.
Contention is a limiting factor regarding scalability: concurrent reads and writes to the same
location must be serialized by the cache coherence protocol, and only a single concurrent \ac{CAS} % TODO: Short section on primitives.
can succeed while all others must retry. \emph{Disjoint-access parallelism} is the concept of
spreading such accesses in order to reduce contention as much as possible.

>>>>

% Synchronization primitives.

Lock-free algorithms and data structures are commonly constructed using synchronization primitives
such as \acf{CAS}, \ac{FAA}, \ac{FAO}, and \ac{TAS}. The \ac{CAS} instruction, which atomically
compares a memory location to an expected value and sets it to a new value if they are equal,
is implemented on most modern architectures and can be considered a basic building block of lock-free
programming. More exotic primitives such as \ac{DCAS} and \ac{DCSS} exist as well, but are not yet
widely available and require inefficient software emulations to be used.

@@

An area in memory accessed frequently by a large number of processes is said to be \emph{contended}.
Contention is a limiting factor regarding scalability: concurrent reads and writes to the same
location must be serialized by the cache coherence protocol, and only a single concurrent \ac{CAS}
can succeed while all others must retry. \emph{Disjoint-access parallelism} is the concept of
spreading such accesses in order to reduce contention as much as possible.
